plyranges: A grammar of genomic data transformation

### 2018-10-23 paper.Rmd
abstract and title change based on editor's comments

---
title: "plyranges: A grammar of genomic data transformation" 
author:
  - name: Stuart Lee \footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}
    email: stuart.lee1@monash.edu
  - name: Dianne Cook $^{\ref{monash}}$
    email: di.cook@monash.edu 
  - name: Michael Lawrence \footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author} 
    email: lawrence.michael@gene.com
abstract: "The Bioconductor project provides many interoperable data abstractions for analyzing high-throughput genomics experiments; however implementing a typical genomic workflow with Bioconductor requires learning these abstractions and understanding them at an integrative level. This places a large cognitive burden on the user, especially for non-programmers. To reduce this burden we have created a grammar of genomic data transformation that operates on a single, central Bioconductor data structure, GRanges, which naturally represents genomic intervals and their associated measurements. The grammar defines verbs for performing actions on and between genomic interval data through a simplified, coherent interface to existing Bioconductor infrastructure, resulting in fluent analysis workflows. We have implemented this grammar as an R/Bioconductor package called \\texttt{plyranges}."
keywords: "Bioconductor, Grammar, Genomes, Data Analysis" 
bibliography: references.bib 
output: 
  pdf_document:
    template: svm-latex-ms.tex 
    keep_tex: true 
    fig_caption: true
    latex_engine: pdflatex 
    citation_package: biblatex 
fontfamily: mathpazo
---

>>>>

---
title: "plyranges: a grammar of genomic data transformation" 
author:
  - name: Stuart Lee \footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}
    email: stuart.lee1@monash.edu
  - name: Dianne Cook $^{\ref{monash}}$
    email: di.cook@monash.edu 
  - name: Michael Lawrence \footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author} 
    email: lawrence.michael@gene.com
abstract: "There is a cognitive load placed on users in learning a data abstraction from the Bioconductor project and understanding its appropriate use. Users must navigate these abstractions to perform a genomic analysis task, when a single data abstraction, a GRanges object will suffice. By recognizing that the GRanges class follows 'tidy' data principles, we create a grammar of genomic data transformation, defining verbs for performing actions on and between genomic interval data and providing a way of performing common data analysis tasks through a coherent interface to existing Bioconductor infrastructure. We implement this grammar as a Bioconductor/R package called \texttt{plyranges}."
keywords: "Bioconductor, Grammar, Genomes, Data Analysis" 
bibliography: references.bib 
output: 
  pdf_document:
    template: svm-latex-ms.tex 
    keep_tex: true 
    fig_caption: true
    latex_engine: pdflatex 
    citation_package: biblatex 
fontfamily: mathpazo
---

### 2018-10-23 paper.tex
abstract and title change based on editor's comments

@@

Recently, there has been a concerted, community effort to standardize R
data structures and workflows around the notion of tidy data
\cite{Wickham2014-jc}. A tidy dataset is defined as a tabular data
structure that has observations as rows and columns as variables, and
all measurements pertain to a single observational unit. The tidy data
pattern is useful because it allows us to see how the data relate to the
design of an experiment and the variables measured. The \texttt{dplyr}
package \cite{Wickham2017-dplyr} defines an application programming
interface (API) that maps notions from the general relational algebra to
verbs that act on tidy data. These verbs can be composed together on one
or more tidy datasets with the pipe operator from the \texttt{magrittr}
package \cite{R-magrittr}. Taken together these features enable a user
to write human readable analysis workflows.

>>>>

% highlight changes
\usepackage{soul}

@@

Recently, there has been a concerted, community effort to standardize R
data structures and workflows around the notion of tidy data
\cite{Wickham2014-jc}. A tidy dataset is defined as a tabular data
structure that has observations as rows and columns as variables, and
all measurements pertain to a single observational unit. The tidy data
pattern is useful because it allows us to see how the data relate to the
design of an experiment and the variables measured. The \texttt{dplyr}
package \cite{Wickham2017-dplyr} defines an application programming
interface (API) that maps notions from query languages such as SQL to
verbs that act on tidy data. These verbs can be composed together on one
or more tidy datasets with the pipe operator from the \texttt{magrittr}
package \cite{R-magrittr}. Taken together these features enable a user
to write human readable analysis workflows.
